---
title: "final project"
author: "Jiaxin Wu"
date: "2023-04-12"
output: html_document
---

```{r setup, include=FALSE}
library(deSolve)
library(tidyverse)
```


### Simulation 1
```{r sim1}
SIRvec <- function(t, state, parameters) {
  with(as.list(c(state, parameters)), {
    
    #Host population
    dSH = NH/TLH - SH*((cVH*IV/NH)+(1/TLH))
    dEH = SH*cVH*(IV/NH) - EH*((1/TIIT)+(1/TLH))
    dIH = EH/TIIT - IH*((1/TID)+(1/TLH))
    dRH = IH/TID - RH/TLH
    
    #Vector population
    dSV = e - SV*(cHV*(IH+IHvisit)/NH + (1/TLV))
    dEV = SV*(cHV*(IH+IHvisit)/NH) - EV*((1/TEIT)+(1/TLV))
    dIV = EV/TEIT - IV/TLV
    
    dcumi = EH/TIIT
    
    list(c(dSH,dEH,dIH,dRH,dSV,dEV,dIV,dcumi))
  })
}


# Initial states and parameters
cHV=0.375
cVH=0.75
TLH=25002.5
TLV=4
TIIT=5
TEIT=10
TID=3
e=5000

SH=10000
EH=0
IH=0
RH=0
SV=19999
EV=0
IV=1
IHvisit=0
NH=SH+EH+IH+RH
NV=SV+EV+IV

parameters = c( cHV=cHV, cVH=cVH,
                TLH=TLH, TLV=TLV,
                TIIT=TIIT, TEIT=TEIT,
                TID=TID, e=5000, IHvisit=IHvisit)

state = c(SH = SH, EH = EH, IH = IH, RH = RH, SV = SV, EV = EV, IV = IV,cumi=0)

times=1:365;

sim1=ode(times=times,func=SIRvec,y=state,parms=parameters)

#prevalence peak without fogging and cumulative incidence 
which.max(sim1[,"IH"])
tail(sim1[,"cumi"],1)

prevalence=matrix(0,364,364)
cumulated_i=matrix(0,364,364)

for (i in 2:364) {
  times1=1:i; 
  times2=i:365; 
  
  simNoCtrl=ode(times=times1,func=SIRvec,y=state,parms=parameters);
  state2=c(SH=tail(simNoCtrl[,'SH'],1),EH=tail(simNoCtrl[,'EH'],1),
           IH=tail(simNoCtrl[,'IH'],1),RH=tail(simNoCtrl[,'RH'],1),
           SV=tail(simNoCtrl[,'SV'],1)*0.4,
           EV=0.4*tail(simNoCtrl[,'EV'],1),
           IV=0.4*tail(simNoCtrl[,'IV'],1),
           cumi=tail(simNoCtrl[,'cumi'],1));
  simCtrl=ode(times=times2,func=SIRvec,y=state2,parms=parameters);
  sim=rbind(simNoCtrl[-1,],simCtrl[-1,])
  prevalence[,i]=sim[,'IH']
  cumulated_i[,i]=sim[,"cumi"]
}

# prevalence peak day with fogging and cumulative incidence 
which.min(tail(cumulated_i[,2:364],1))
min(tail(cumulated_i[,2:364],1))
# 7121.874 vs Table 2 7120. => similar


matplot(sim1[,'time'], sim1[,"IH"], type='l',xlim = c(0,365),
        lwd = 1, lty = 1, cex.main=1,
        ylab = 'Prevalence',xlab = 'Time(days)')
lines(1:364, y=prevalence[,168], type = 'l',col="red")
legend("topright",legend=c('without fogging', 'with fogging'),col=c('black', 'red'), lty=1)

```


### simulation 2
```{r sim2}
times <- 1:365

wetseason=c(1:150);  # the days in a year that are wetseason
Term=rep(3,length(times));  # initialize a vector to store the Term
ind=(1:length(Term) %% 365) %in% wetseason  # find those days that are wetseason
Term[ind]=4; # set them to 4


# Modify SIRvec function
SIRseason <- function(time, state, parameters) {
  with(as.list(c(state, parameters)), {
    
    #TLV value changes based on time 
    TLV = Term[time]
        
    #Host population
    dSH = NH/TLH - SH*((cVH*IV/NH)+(1/TLH))
    dEH = SH*cVH*(IV/NH) - EH*((1/TIIT)+(1/TLH))
    dIH = EH/TIIT - IH*((1/TID)+(1/TLH))
    dRH = IH/TID - RH/TLH
    
    #Vector population
    dSV = e - SV*(cHV*(IH+IHvisit)/NH + (1/TLV))
    dEV = SV*(cHV*(IH+IHvisit)/NH) - EV*((1/TEIT)+(1/TLV))
    dIV = EV/TEIT - IV/TLV
    
    dcumi = EH/TIIT
    
    list(c(dSH,dEH,dIH,dRH,dSV,dEV,dIV,dcumi))
  })
}


cHV=0.375
cVH=0.75
TLH=25002.5


TIIT=5
TEIT=10
TID=3
e=5000

SH=10000
EH=0
IH=0
RH=0
SV=19999
EV=0
IV=1
IHvisit=0
NH=SH+EH+IH+RH
NV=SV+EV+IV

parameters <- c(cHV = cHV, cVH = cVH,
                TLH = TLH, TLV_wet = TLV_wet, TLV_dry = TLV_dry,TIIT = TIIT, TEIT = TEIT,TID = TID, e = 5000, IHvisit = IHvisit)

state = c(SH = SH, EH = EH, IH = IH, RH = RH, SV = SV, EV = EV, IV = IV,cumi=0)


sim2=ode(times=times,func=SIRseason,y=state,parms=parameters)

which.max(sim2[,"IH"])
tail(sim2[,"cumi"],1)

prevalence=matrix(0,364,364)
cumulated_i=matrix(0,364,364)

for (i in 2:364) {
  times1=1:i; 
  times2=i:365;
  
  simNoCtrl=ode(times=times1,func=SIRseason,y=state,parms=parameters);
  state2=c(SH=tail(simNoCtrl[,'SH'],1),EH=tail(simNoCtrl[,'EH'],1),
           IH=tail(simNoCtrl[,'IH'],1),RH=tail(simNoCtrl[,'RH'],1),
           SV=tail(simNoCtrl[,'SV'],1)*0.4,
           EV=0.4*tail(simNoCtrl[,'EV'],1),
           IV=0.4*tail(simNoCtrl[,'IV'],1),
           cumi=tail(simNoCtrl[,'cumi'],1));
  simCtrl=ode(times=times2,func=SIRseason,y=state2,parms=parameters);
  sim=rbind(simNoCtrl[-1,],simCtrl[-1,])
  
  prevalence[,i]=sim[,'IH']
  cumulated_i[,i]=sim[,"cumi"]
}

# prevalence peak day with fogging and cumulative incidence 
which.min(tail(cumulated_i[,6:364],1))
tail(cumulated_i[,59],1)

matplot(sim2[,'time'], sim2[,"IH"], type='l',xlim = c(0,365),
        lwd = 1, lty = 1, cex.main=1,
        ylab = 'Prevalence',xlab = 'Time(days)')
lines(1:364, y=prevalence[,108], type = 'l',col="red")
legend("topright",legend=c('without fogging', 'with fogging'),col=c('black', 'red'), lty=1)


```

### Simulation 3 (part 1)
#### before year 500

```{r sim3}
wetseason=c(1:150);  # the days in a year that are wetseason
times=seq(1,499*365); # in day
Term=rep(3,length(times));  # initialize a vector to store the Term
ind=(1:length(Term) %% 365) %in% wetseason  # find those days that are wetseason
Term[ind]=4; # set them to 4

# Modify SIRvec function
SIRseason <- function(time, state, parameters) {
  with(as.list(c(state, parameters)), {

    #TLV value changes based on time 
    TLV = Term[time]
    
    #Host population
    dSH = NH/TLH - SH*((cVH*IV/NH)+(1/TLH))
    dEH = SH*cVH*(IV/NH) - EH*((1/TIIT)+(1/TLH))
    dIH = EH/TIIT - IH*((1/TID)+(1/TLH))
    dRH = IH/TID - RH/TLH
    
    #Vector population
    dSV = e - SV*(cHV*(IH+IHvisit)/NH + (1/TLV))
    dEV = SV*(cHV*(IH+IHvisit)/NH) - EV*((1/TEIT)+(1/TLV))
    dIV = EV/TEIT - IV/TLV
    
    dcumi = EH/TIIT
    
    list(c(dSH,dEH,dIH,dRH,dSV,dEV,dIV,dcumi))
  })
}


cHV=0.375
cVH=0.75
TLH=25002.5

TIIT=5
TEIT=10
TID=3
e=5000

SH=10000
EH=0
IH=0
RH=0
SV=19999
EV=0
IV=1
IHvisit=0.001
NH=SH+EH+IH+RH
NV=SV+EV+IV

parameters <- c(cHV = cHV, cVH = cVH,
                TLH = TLH,TIIT = TIIT, TEIT = TEIT,TID = TID, e = 5000, IHvisit = IHvisit)

state = c(SH = SH, EH = EH, IH = IH, RH = RH, SV = SV, EV = EV, IV = IV, cumi=0)

sim3=ode(times=times,func=SIRseason,y=state,parms=parameters)
plot(sim3[,"time"],sim3[,"IH"],xlab="Days",ylab="Prevalence", type = "l", ylim = c(0,3),xlim = c(498*365,499*365))

```

### simulation 3 (part 2)
#### year 500 without fogging

```{r}
wetseason=c(1:150);  # the days in a year that are wetseason
times=seq(1,365); # in day
Term=rep(3,length(times));  # initialize a vector to store the Term
ind=(1:length(Term) %% 365) %in% wetseason  # find those days that are wetseason
Term[ind]=4; # set them to -1

SIRseason <- function(time, state, parameters) {
  with(as.list(c(state, parameters)), {
    
    #TLV value changes based on time 
    TLV = Term[time]
    
    #Host population
    dSH = NH/TLH - SH*((cVH*IV/NH)+(1/TLH))
    dEH = SH*cVH*(IV/NH) - EH*((1/TIIT)+(1/TLH))
    dIH = EH/TIIT - IH*((1/TID)+(1/TLH))
    dRH = IH/TID - RH/TLH
    
    #Vector population
    dSV = e - SV*(cHV*(IH+IHvisit)/NH + (1/TLV))
    dEV = SV*(cHV*(IH+IHvisit)/NH) - EV*((1/TEIT)+(1/TLV))
    dIV = EV/TEIT - IV/TLV
    
    dcumi = EH/TIIT
    
    list(c(dSH,dEH,dIH,dRH,dSV,dEV,dIV,dcumi))
  })
}



parameters <- c(cHV = cHV, cVH = cVH,
                TLH = TLH, TIIT = TIIT, TEIT = TEIT,TID = TID, e = 5000, IHvisit = IHvisit)

  state1=c(SH=tail(sim3[,'SH'],1),EH=tail(sim3[,'EH'],1),
           IH=tail(sim3[,'IH'],1),RH=tail(sim3[,'RH'],1),
           SV=tail(sim3[,'SV'],1),EV=tail(sim3[,'EV'],1),
           IV=tail(sim3[,'IV'],1),cumi=0);


sim4=ode(times=times,func=SIRseason,y=state1,parms=parameters)

#prevalence peak without fogging 
which.max(sim4[,"IH"])
#156

#cumulative incidence without fogging 
tail(sim4[,"cumi"],1)
#28.63
  
plot(sim4[,"time"],sim4[,"IH"],xlab="Days",ylab="Prevalence", type = "l", ylim = c(0,3))

#head imuunity cuculation 
herd = 1-tail(sim4[,"SH"],1)/NH
```

### simulation 3 (part 2)
#### year 500 with fogging

```{r}
prevalence=matrix(0,365,364)
infected=matrix(0,365,364)
cumulated_i=matrix(0,365,364)

for (i in 2:364) {
  times1=1:i; 
  times2=i:365;

  simNoCtrl=ode(times=times1,func=SIRseason,y=state1,parms=parameters);
  state2=c(SH=tail(simNoCtrl[,'SH'],1),EH=tail(simNoCtrl[,'EH'],1),
           IH=tail(simNoCtrl[,'IH'],1),RH=tail(simNoCtrl[,'RH'],1),
           SV=tail(simNoCtrl[,'SV'],1)*0.4,
           EV=0.4*tail(simNoCtrl[,'EV'],1),
           IV=0.4*tail(simNoCtrl[,'IV'],1),
           cumi=tail(simNoCtrl[,"cumi"],1));
  simCtrl=ode(times=times2,func=SIRseason,y=state2,parms=parameters);
  sim=rbind(simNoCtrl,simCtrl[-1,])
  
  prevalence[,i]=sim[,'IH']
  infected[,i]=sim[,"RH"]+sim[,"IH"]
  cumulated_i[,i]=sim[,"cumi"]
}


which.min(tail(infected[,2:364],1))
#56 day fogging produce the least infected population
min(tail(infected[,2:364],1))
#the number of least infected = 3848.951

which.min(tail(cumulated_i[,2:364],1))
#63 day fogging produce the least infected population
min(tail(cumulated_i[,2:364],1))
#the number of least infected = 16.92

matplot(sim4[,"time"], sim4[,"IH"], type='l',
        lwd = 1, lty = 1, cex.main=1,ylim = c(0,3),
        ylab = 'Prevalence',xlab = 'Time(days)')
lines(1:365, y=prevalence[,18], type = 'l',col="red")
legend("topright",legend=c('without fogging', 'with fogging'),col=c('black', 'red'), lty=1)


```



# simulation 4: 
## before year 500
```{r}
SH=10000
EH=0
IH=0
RH=0
NH=SH+EH+IH+RH

NV=c(20000,30000,50000,80000,150000)
SV=c(19999,29999,49999,79999,149999)
IV=1
EV=RV=0

TLH=25002.5
TIIT=5
TEIT=10
e=c(5000,7500,12500,20000,37500)
TLV_dry=3
TLV_wet=4
IHvisit=0.001
TID=3
cVH=0.75
cHV=0.375


wetseason=c(1:180);  # the days in a year that are wetseason
times=seq(1,499*365); # in day
Term=rep(3,length(times));  # initialize a vector to store the Term
ind=(1:length(Term) %% 365) %in% wetseason  # find those days that are wetseason
Term[ind]=4; # set them to 4

SIRseason <- function(time, state, parameters) {
  with(as.list(c(state, parameters)), {
    
    #TLV value changes based on time 
    TLV = Term[time]
    
    #Host population
    dSH = NH/TLH - SH*((cVH*IV/NH)+(1/TLH))
    dEH = SH*cVH*(IV/NH) - EH*((1/TIIT)+(1/TLH))
    dIH = EH/TIIT - IH*((1/TID)+(1/TLH))
    dRH = IH/TID - RH/TLH
    
    #Vector population
    dSV = e - SV*(cHV*(IH+IHvisit)/NH + (1/TLV))
    dEV = SV*(cHV*(IH+IHvisit)/NH) - EV*((1/TEIT)+(1/TLV))
    dIV = EV/TEIT - IV/TLV
    
    dcumi = EH/TIIT
    
    list(c(dSH,dEH,dIH,dRH,dSV,dEV,dIV,dcumi))
  })
}

state_500=matrix(0,5,9)


for (i in 1:5) {
  parameters = c( cHV=cHV, cVH=cVH,
                TLH=TLH, TLV_wet = TLV_wet, TLV_dry = TLV_dry,
                TIIT=TIIT, TEIT=TEIT,
                TID=TID, e=e[i], IHvisit=IHvisit,NV=NV[i])

state = c(SH = SH, EH = EH, IH = IH, RH = RH, SV = SV[i], EV = EV, IV = IV,dcumi=0)

sim4=ode(times=times,func=SIRseason,y=state,parms=parameters)

state_500[i,]=tail(sim4[],1)

}

colnames(state_500)<-c("end","SH","EH","IH","RH","SV","EV","IV","cumi")
rownames(state_500)<-c("2","3","5","8","15")

herd = 1-state_500[,"SH"]/NH
```

###"state_500" is the initial states for the beginning of year 500 for each MPP


## year 500 without fogging
```{r}
#change wetseason
wetseason=c(1:180);
times=seq(1,365)
infected=matrix(0,365,5)
prevalence=matrix(0,365,5)
peak_without=matrix(c(2,3,5,8,15),5,2)
cumi500=matrix(c(2,3,5,8,15),5,2)

for (i in 1:5) {
  parameters <- c(cHV = cHV, cVH = cVH,
                TLH = TLH, TLV_wet = TLV_wet, TLV_dry = TLV_dry,TIIT = TIIT, TEIT = TEIT,TID = TID, e = e[i], IHvisit = IHvisit)

  state500=c(SH=state_500[i,2],EH=state_500[i,3],
           IH=state_500[i,4],RH=state_500[i,5],
           SV=state_500[i,6],EV=state_500[i,7],
           IV=state_500[i,8],cumi=0);
  
  sim500_0=ode(times=times,func=SIRseason,y=state500,parms=parameters)
  prevalence[,i]=sim500_0[,'IH']
  infected[,i]=sim500_0[,"RH"]+sim500_0[,"IH"]
  peak_without[i,2]=which.max(prevalence[,i])
  cumi500[i,2]=tail(sim500_0[,'cumi'],1)
}

which.max(prevalence[,2])

```

### "cumi500" is the cumulative incidence without fogging for year 500


## year 500 with fogging
```{r}
cumulated_i=data.frame()
out=data.frame()

for (j in 1:5) {
  
  state500=c(SH=state_500[j,2],EH=state_500[j,3],
           IH=state_500[j,4],RH=state_500[j,5],
           SV=state_500[j,6],EV=state_500[j,7],
           IV=state_500[j,8],cumi=0);
  
    parameters <- c(cHV = cHV, cVH = cVH,
                TLH = TLH, TLV_wet = TLV_wet, TLV_dry = TLV_dry,TIIT = TIIT, TEIT = TEIT,TID = TID, e = e[j], IHvisit = IHvisit)
    
    cum_0=cumi500[j,2]
  
  for (i in 2:364) {
    times1=1:i; 
    times2=i:365;

  simNoCtrl=ode(times=times1,func=SIRseason,y=state500,parms=parameters);
  
  state2=c(SH=tail(simNoCtrl[,2],1),EH=tail(simNoCtrl[,3],1),
           IH=tail(simNoCtrl[,4],1),RH=tail(simNoCtrl[,5],1),
           SV=tail(simNoCtrl[,6],1)*0.4,
           EV=0.4*tail(simNoCtrl[,7],1),
           IV=0.4*tail(simNoCtrl[,8],1),
           cumi=tail(simNoCtrl[,9],1));
  
  simCtrl=ode(times=times2,func=SIRseason,y=state2,parms=parameters);
  
  sim=rbind(simNoCtrl,simCtrl[-1,])
  cumulated_i[(i-1),j]=tail(sim[,"cumi"],1)
  out[(i-1),j]=(cum_0-tail(sim[,"cumi"],1))/cum_0

  }
  
    
}

out=out %>% 
  mutate(time = 2:364) 

colnames(out)=c("2","3","5","8","15","time")

which.max(out[,5])+1
cumulated_i[75,5]


```

### "out" is the percent of prevented cases for each MPP, "time" in "out" is the intervention day.

